A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates

### 2019-08-02 paper-submitted-2.tex
Updates from Overleaf

One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^{\circ}$ N, geodetic longitude: 144.189$^{\circ}$ E) were obtained from the Japan Meteorological Agency data portal. The location of the observatory and substation are shown in Figure~\ref{map}.

@@

\noindent
where, explicitly, the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.

@@

\noindent
In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.

Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data. Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used. The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}. The regression for each $f_e$ uses all DFT frequencies with values in a band with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest band has a lower bound of $1/86400$~Hz and highest band has an upper bound of $0.375$~Hz.

The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.

For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using

@@

\begin{linenomath*}
  \begin{equation*}
    E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)
  \end{equation*}
\end{linenomath*}

Transfer functions were computed for each of the 1-day segments. The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.

@@

GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(\omega)=a^I(\omega)E_x(\omega)+b^I(\omega)E_y(\mega)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(\omega)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).

Several modifications to this basic procedure were considered. Robust regression \citep{Egbert1986} with either a Huber weighting using a constant of $1.345$ or bi-square weighting using a constant of $4.685$ (both with no hard-cutoff step). The commonly used constants provide 95\% efficiency for gaussian-distributed errors. Models were also computed using time domain pre-whitening (using a 10th-order Yule-Walker model computed using the model output), time domain tapering (using a Parzen window), and weighting the frequencies around each evaluation frequency using Parzen window weights. When using these additional procedures individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method. Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal-to-noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s); at periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)

@@

The mean-squared-error for model $m$ on day $D$ is
\begin{equation*}
  \mbox{MSE}_m = \frac{1}{N}\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2 \,,
\end{equation*}
\noindent
where $GIC_m(t)$ is the estimate from model $m$ at time step $t$;  $t_o=600$ and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes. The MSE ratio for model $m$ is MSE$_1$/MSE$_m$.

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is calculated for day $D$ using
\begin{equation*}
  \mbox{PE}_m = 1 - \frac{\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2}{\sum_{t=t_o}^{t_f} \big(GIC(t)-\overline{GIC(t)}\big)^2}\,,
\end{equation*}
\noindent
where $\overline{GIC(t)}$ is an average over the same time interval used in the summations. The PE is a measure of the skill of the model with respect to the variance in the predictand and it represents the fraction of the variance in the data that is predicted by the model (when PE~$\ge 0)$.

@@

The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency. A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}. The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.

@@

\begin{table}
  \caption{Out-of-sample model performance metrics for Model $m$. MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
  \centering
  \begin{tabular}{l c c c c}
    \hline
    $m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
    \hline
    1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
    2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
    3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
    4\hspace{1em} $G_{B}(t) = z_xB_x(t) + z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
    \hline
  \end{tabular}
  \label{resultstable}
\end{table}

@@

Below $600$~s, Model~2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models. For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a larger SN ratio than Model~3, which uses an indirect estimate of the electric field. Model~4 has a slightly larger SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.

Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field. Issues related to this are discussed in the following section.

@@

The average coefficients and 95\% CI for Model 1 are $\overlin{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14 to 2006-12-15 JST time interval).

The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction. \cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.

@@

Figure~\ref{Phi} shows the transfer function coefficient phase angles (that for Model~1 is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition). The phases for $a(\omega)$ and $b(\omega)$ are frequency dependent. At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180$^{\circ}$ and the associated signs are consistent with those of $\overline{a}_o > 0$ and $\overline{b}_o < 0$. At $T=12$~h, the values of $b(\omega)$ and $a(\omega)$ and Model~2 correspond to an electric field of 60$^{\circ}$ west of north being associated with the largest GIC.

Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line. As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity. 
Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$. That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations. After 2006 (January through May tested) our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work. This result suggests that stack averaging can be a significant source of transfer function bias and we expect that the explanation involves variations in the signal-to-noise between segments.

@@

\cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency. Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower. Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance. (However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant of the power network may be large enough to warrant inclusion in GIC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)

(2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence. However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (Oowada et al. 1998, Fujii 2019 personal communication.)

(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems. (An equation of this form can be computed in principle for an electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform). The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}. The voltages are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system. In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field. Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces. If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information. However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of the electric field along the line.

@@

To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:

@@

Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated. (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}). However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.

@@

When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used. For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE and MSE metrics. Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions. A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and more readily available than geoelectric field measurements. A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/map.pdf}
  \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memanbetsu Magnetic Observatory. The red line corresponds to the path of the 187 kV power line. The distance from the observatory to the substation is $\sim$9 km and dotted lines are Apex geomagnetic latitudes \citep{Richmond1995} in 2015.}
  \label{map}
\end{figure}

>>>>

One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^{\circ}$ N, geodetic longitude: 144.189$^{\circ}$ E) were obtained from the Japan Meteorological Agency data portal. The locations of the observatory and substations are shown in Figure~\ref{map}.

@@

\noindent
where, explicitly, the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
That is, instead of using the measured electric field directly as in Model~2, a transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.

@@

\noindent
In this model, the transfer function components $z_x(\omega)$ and $z_y(\omega)$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.

Estimates of $a(\omega)$ and $b(\omega)$ in Model~2 are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS). At each evaluation frequency, a regression is performed using a set $G_E(\omega)$, $E_x(\omega)$, and $E_y(\omega)$ values from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data in a band of frequencies around $\omega_e$. Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used. The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}. The regression for each $f_e$ uses all DFT frequencies with values in a band with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest band has a lower bound of $1/86400$~Hz and highest band has an upper bound of $0.375$~Hz.

The same approach is used for solving for $z_x(\omega_e)$ and $z_y(\omega_e)$ in Equation~\ref{model4}.

For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ at the evaluation frequencies are solved for using the same method used for Model~2 using the equation

@@

\begin{linenomath*}
  \begin{equation*}
    E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)\,.
  \end{equation*}
\end{linenomath*}

Transfer functions were computed for each of the 1-day segments. The average transfer function magnitude and phase at a given evaluation frequency were obtained by averaging the segment values at that evaluation frequency.

@@

GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(\omega)=a^I(\omega)E_x(\omega)+b^I(\omega)E_y(\omega)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(\omega)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).

Several modifications to this basic procedure were considered. Robust regression \citep{Egbert1986} with either a Huber weighting using a constant of $1.345$ or bi-square weighting using a constant of $4.685$, both with no hard-cutoff step. The commonly used constants provide 95\% efficiency for Gaussian-distributed errors. Models were also computed using time domain pre-whitening (using a 10th-order Yule-Walker model computed using the model output), time domain tapering (using a Parzen window), and weighting the frequencies around each evaluation frequency using Parzen window weights. When using these additional procedures individually, we find that the conclusions made regarding the differences in each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method. Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal-to-noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s); at periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.

@@

The mean-squared-error for model $m$ is
\begin{equation*}
  \mbox{MSE}_m = \frac{1}{N}\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2 \,,
\end{equation*}
\noindent
where $GIC_m(t)$ is the estimate from model $m$ at time step $t$;  $t_o=600$ and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes. The MSE ratio for model $m$ is MSE$_1$/MSE$_m$.

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is calculated using
\begin{equation*}
  \mbox{PE}_m = 1 - \frac{\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2}{\sum_{t=t_o}^{t_f} \big(GIC(t)-\overline{GIC(t)}\big)^2}\,,
\end{equation*}
\noindent
where $\overline{GIC(t)}$ is an average over the same time interval used in the summations. The PE is a measure of the skill of the model with respect to the variance in the predictand and it represents the fraction of the variance in the data that is predicted by the model (when PE~$\ge 0)$.

@@

The frequency dependent signal-to-noise ratio for each 1-day segment at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency. A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}. The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.

@@

\begin{table}
  \caption{Out-of-sample model performance metrics for Model $m$. MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
  \centering
  \begin{tabular}{l c c c c}
    \hline
    $m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
    \hline
    2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
    3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
    4\hspace{1em} $G_{B}(t) = z_xB_x(t) + z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
    \hline
  \end{tabular}
  \label{resultstable}
\end{table}

@@

Below $600$~s, Model~2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models. Below $600$~s, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a larger SN ratio than Model~3, which uses an indirect estimate of the electric field. Model~4 has a slightly larger SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.

Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) the error associated with this assumption has less of an influence on the errors in a model that uses point measurements of the magnetic field. Issues related to this are discussed in the following section.

@@

The average coefficients and 95\% CI for Model 1 are $\overline{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14 to 2006-12-15 JST time interval).

The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points 45$^{\circ}$ west of north, which is approximately perpendicular to the 187~kV line shown in Figure~\ref{map} that runs in the northeast direction. \cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to a maximum GIC when the electric field angle is 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.

@@

Figure~\ref{Phi} shows the transfer function coefficient phase angles. For Model~1, the phase angle is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition. The phases for $a(\omega)$ and $b(\omega)$ are frequency dependent. At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180$^{\circ}$ and the associated signs are consistent with those of $\overline{a}_o > 0$ and $\overline{b}_o < 0$. At $T=12$~h, the values of $b(\omega)$ and $a(\omega)$ and Model~2 correspond to an electric field of 60$^{\circ}$ west of north being associated with the largest GIC.

Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line. As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and differences in ground conductivity in parts of the network that contribute to GIC.

Figure~\ref{Z_MT} and~\ref{Phi_MT} show the magnitude and phase for the coefficients of $\mathbf{Z}$ in Model~3. Note that transfer function coefficients shown in Figure~\ref{Z_MT} differ significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$. That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function coefficients match the \cite{Fujii2015} coefficients to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations. However, using one-month intervals after 2006 (January through May tested) our calculated transfer function coefficient magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer functions were computed using one-day segments of data and then averaged. We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function coefficient magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged. This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work. This result suggests that stack averaging can be a significant source of transfer function bias and we expect that the explanation involves variations in the signal-to-noise between segments.

@@

\cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency. Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower. Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth to more accurately model skin an proximity effects also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance. (However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant of the power network may be large enough to warrant inclusion in GIC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)

(2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence. However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (\cite{Oowada1998}; Fujii 2019 personal communication.)

(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems. (An equation of this form can be computed in principle for a static electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform). The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}. The electric potential differences are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system. In both Model~1 and Model~2, the input electric field is typically assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field. Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces. If a spatial variation exists along the line in which GIC was measured, the coefficients derived from measurements in both models may differ significantly from those calculated using system configuration and resistance information. However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of a static electric field along the line.

@@

To test the effect of different points on the line having different electric field spectra on the coefficient estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:

@@

Although the use of frequency independent $a$ and $b$ coefficients estimated from data give a reasonable estimates of GIC, it is not known how much data-estimated coefficient values will differ from those directly calculated. (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}). However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.

@@

When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used. For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE, MSE, and signal-to-error metrics. Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions. A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and more readily available than geoelectric field measurements. A disadvantage is the transfer function coefficients of Model~4 are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/map.pdf}
  \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memanbetsu Magnetic Observatory. The red line corresponds to the path of the 187 kV power line. The distance from the observatory to the Memanbetsu substation is $\sim$9 km and dotted lines are Apex geomagnetic latitudes \citep{Richmond1995} in 2015.}
  \label{map}
\end{figure}
